Pure Spin Current Injection in Hydrogenated Graphene Structures

### 2017-06-11 mu/article/article.tex
Completed intro redaction of Spin Velocity Injection.

\usepackage{amsmath}    % need for subequations
\usepackage{graphicx}   % need for figures
\usepackage{verbatim}   % useful for program listings
\usepackage{color}      % use if color is used in text
\usepackage{subfigure}  % use for side-by-side figures
\usepackage{hyperref}   % use for hypertext links, including those to external
                        % documents and URLs 
\usepackage{blindtext}  % fill text
\usepackage[dvipsnames]{xcolor}
\usepackage{multirow}
% \usepackage{showframe}
\usepackage[inline]{showlabels}

@@

\section{Introduction}

@@

\subsection{Pure spin velocity} % (fold)
\label{sec:theory-pure_spin_current}

The spin density injection current $\dot{K}^{\mathrm{ab}}(\omega)$, 
% 
{
\color{red}
where the roman superscripts denote Cartesian components corresponding
$\mathrm{a}$ to the spin velocity direction and $\mathrm{b}$ to the spin
polarization direction, can be written in terms of a material response
pseudotensor $\mu^{\mathrm{abcd}}(\omega)$ as }
\begin{equation*}
\dot{K}^{\mathrm{ab}}(\omega) = \mu^{\mathrm{abcd}}(\omega)
E^{\mathrm{c}}(\omega) E^{\mathrm{d*}}(\omega)
\label{eq:dotk}
\end{equation*}
where the pseudotensor is given by\cite{bhatPRL05}
\begin{widetext}
\begin{align}
\mu^{\mathrm{abcd}} (\omega) 
&=
\frac{\pi e^{2}}{\hbar^{2}} \int 
\frac{d^{3}K}{8 \pi^{3}}
\sum_{vcc'}^{'}
\mathrm{Re} \left[ K^{\mathrm{ab}}_{cc'} 
\left( 
r^{\mathrm{c}}_{vc'} 
r^{\mathrm{d}}_{cv } +
(c \leftrightarrow d)
\right) 
\right]
\delta(\omega-\omega_{cv})
\label{eq:mu}
\end{align}
\end{widetext}
{\color{red}
where $K^{\mathrm{ab}}_{mn}(\mathbf{k}) =
\sum_{\ell}v^{\mathrm{a}}_{nl}(\mathbf{k}) S^{\mathrm{b}}_{lm}(\mathbf{k})$ are
the spin current matrix elements that, using time reversal invariance, satisfy
the relation $K^{\mathrm{ab}}_{nm}(\mathbf{-k}) = K^{\mathrm{ab*}}(\mathbf{k})
$. 
}% 
The $'$ in the sum means that $c$ and $c'$ are quasi degenerate states and
the sum only covers these states and 
{\color{red}
since $\mu$ is real we have that
$\mu^{\mathrm{abcd}}(\omega) = \mu^{\mathrm{abdc}}(\omega)$.
}
Now we define the spin velocity, $\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ as
the speed at which the spin polarized in the $\mathrm{b}$   direction moves
along the $\mathrm{a}$ direction when a normal incident beam reaches the $xy$
plane with a polarization angle $\alpha$ as
\begin{widetext}
\begin{align}
\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)
&= \frac{2}{\hbar}
\frac{\mu^{\mathrm{abxx}}(\omega)
E^{2}(\omega)\cos^{2}(\alpha) + 
\mu^{\mathrm{abyy}}(\omega)
E^{2}(\omega)\sin^{2}(\alpha) + 
2\mu^{\mathrm{abxy}}(\omega)
E^{2}(\omega)\cos(\alpha)\sin(\alpha)}
{\xi^{\mathrm{xx}}(\omega)
E^{2}(\omega)\cos^{2}(\alpha) + 
\xi^{\mathrm{yy}}(\omega)
E^{2}(\omega)\sin^{2}(\alpha)},
\nonumber \\
&= \frac{2}{\hbar}
\frac{\mu^{\mathrm{abxx}}(\omega)\cos^{2}(\alpha) + 
\mu^{\mathrm{abyy}}(\omega)\sin^{2}(\alpha) + 
\mu^{\mathrm{abxy}}(\omega)\sin(2\alpha)}
{\xi^{\mathrm{xx}}(\omega)\cos^{2}(\alpha) + 
\xi^{\mathrm{yy}}(\omega)\sin^{2}(\alpha)},
\label{eq:vab}
\end{align}
\end{widetext}
{\color{red}
where $\xi^{\mathrm{aa}}(\omega)$ are the carrier generation rate tensor
components\cite{arzatePRB14}.
}
% 
For an angle $\alpha = \frac{\pi}{4}$ this expression can be reduced to 
\begin{equation*}\label{eq:vab-90deg}
\mathcal{V}^{\mathrm{ab}} (\omega)
= \frac{2}{\hbar}
\frac{\mu^{\mathrm{abxx}}(\omega) + \mu^{\mathrm{abyy}}(\omega) + 
2\mu^{\mathrm{abxy}}(\omega)}
{\xi^{\mathrm{xx}}(\omega) + \xi^{\mathrm{yy}}(\omega)}.
\end{equation*}

@@

{\color{red} Considering that we have 2D structures, one of the options to
analyze the spin velocity (Eq. \eqref{eq:vab}) is fixing the spin parallel to
the surface of the structure in the $x$ or $y$ directions or perpendicularly to
it in the $z$ direction. We define the magnitude of the spin velocity with spin
polarization along the $\mathrm{b}$ Cartesian coordinate as}
\begin{align}
|\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)| 
&=
\sqrt{
[\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)]^{2}\ +
[\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)]^{2}\ 
}, 
\label{eq:vs-mag}
\end{align}
and the angle at which the spin velocity is directed on the $xy$ plane as
\begin{align}
\gamma_{\mathrm{b}} (\omega,\alpha)
&=
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)} \right),
\label{eq:gamma-ang}
\end{align}
where the angle is measured in the counter-clockwise direction from the
positive $x$ Cartesian coordinate. {\color{red} We also define two special
angles
\begin{align}
\gamma_{\mathrm{b \parallel}}(\omega,\alpha) &= \alpha, 
\label{eq:gamma-par} 
\\
\gamma_{\mathrm{b \perp}}(\omega,\alpha) &= \alpha \pm 90^{\circ}.
\label{eq:gamma-perp}
\end{align}
The first corresponds to the case when the spin velocity is directed, on the
$xy$ plane in the same direction of the polarization angle of the incoming
beam; the second one corresponds to the case when the spin velocity is directed
perpendicularly respect to the polarization angle of the incoming beam. }

@@

{\color{red} Another possibility to analyze the Eq. \eqref{eq:vab} is fixing
the velocity on the $xy$ plane along $x$ or $y$ Cartesian coordinate and then
define the magnitude of the spin velocity directed along the $\mathrm{a}$
direction as}
\begin{align}
|\mathcal{V}^{\mathrm{a}}&(\omega,\alpha)| = \nonumber \\
&\sqrt { 
[\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{az}}(\omega,\alpha)]^{2} .
},
\label{eq:vv-mag}
\end{align}
{\color{red} Then, the spin direction depends of the $x$, $y$, and $z$
components of the previous equation and so we define the spin orientation polar
and azimuthal angles as}
\begin{align}
\theta_{\mathrm{a}}  (\omega,\alpha)
=& 
\cos^{-1} \left( \frac{\mathcal{V}^{\mathrm{az}}(\omega,\alpha)}
{|\mathcal{V}^{\mathrm{a}}(\omega,\alpha)|} \right),
& 0 \leq &\theta \leq \pi, 
\label{eq:polar-ang}
\\
\varphi_{\mathrm{a}} (\omega,\alpha)
=& 
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)} \right),
& 0 \leq &\varphi \leq 2\pi.
\label{eq:azimuthal-ang} 
\end{align}
{\color{red} where $\theta_{\mathrm{a}}(\omega,\alpha)$ is measured from the
positive to the negative $z$ Cartesian coordinate and
$\varphi_{\mathrm{a}}(\omega,\alpha)$ is measured on the $xy$ plane in the
counter-clockwise direction from the positive $x$ Cartesian coordinate.}

@@

\subsection{Spin velocity} % (fold)
\label{sec:res-spin_velocity}

@@

{\color{red}Using the Eq. \eqref{eq:vab}, we calculated the
$\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ response for the \emph{up} and
\emph{alt} 2D structures and for the CdSe and GaAs bulk systems; the results
are presented in Fig. \ref{fig:vab-str-comp}. The angle $\alpha$ presented in
the response of each structure is that for which the response is maximized in
each case.}
% 
From the figure we have that the onset of the response starts when the energy
of the incoming beam is the same of the gap energy.
% 
The most intense response corresponds to the \emph{up} structure centered at
0.088\,eV corresponding to the {\color{red} Mid Infrared (MIR)} radiation and
reaching a spin velocity of 87.2\,Km/s.
% 
In the other hand, for an energy range from 0.66\,eV to 3.0\,eV, corresponding
to energies of the {\color{red} Near Infrared (NIR) to visible radiation}, all
the four structures have contributions in the same order of magnitude.
% 
For this energy range the \emph{up} structure has two peaks centered at
1.94\,eV and 1.97\,eV reaching spin velocities of 22.2\,Km/s and -29.7\,Km/s,
respectively, and the \emph{alt} structure has two peaks centered at 0.72\,eV
and 0.91\,eV reaching spin velocities of -40.2\,Km/s and -32.9\,Km/s,
respectively.
% 
Then, for the bulk structures we have that the CdSe has only one intense
response centered at 0.91\,eV reaching a spin velocity of -26.9\,Km/s, and the
GaAs structure has a large and almost planar zone where the response is held
reaching the maximum for an incoming beam of energy of 2.31\,eV and resulting
in a spin velocity of -21.6\,Km/s.
% 
{\color{red} A negative quantity in the spin velocity means a change in the
spin polarization traveling in the opposite direction.
% 
In table \ref{tab:vab-str-comp} we present the comparison of this values for
the 2D and bulk structures. We found that the most intense response for the
spin velocity corresponds to the \emph{up} structure being 3.25 times more
intense than that of the CdSe and 4.03 times more intense than that of the GaAs
bulk structures. Also, the \emph{alt} structure has a response more intense
than the bulk systems but being less intense than the corresponding to the
\emph{up} one.}
%

>>>>

\usepackage{amsmath}    % need for subequations
\usepackage{graphicx}   % need for figures
\usepackage{verbatim}   % useful for program listings
\usepackage{color}      % use if color is used in text
\usepackage{subfigure}  % use for side-by-side figures
\usepackage{hyperref}   % use for hypertext links, including those to external
                        % documents and URLs 
\usepackage{blindtext}  % fill text
\usepackage[dvipsnames]{xcolor}
\usepackage{multirow}
\usepackage{showframe}
\usepackage[inline]{showlabels}

@@

\section{Introduction}
\label{sec:introduction}

@@

In this section, we report a summary of the theory that involves the PSC
phenomena from which rises the SVI treated in this paper. The full description 
of it was presented by Bhat and Sipe et. al.\cite{bhatPRL05}

As mentioned before in Section \ref{sec:introduction}, In PSCs there is no net
motion of charge and spin-up electrons move in a given direction while spin-
down electrons travel in the opposite one. This effect can result from one
photon absorption of linearly polarized light by a semiconductor, with filled
valence bands and empty conduction bands, illuminated by light with photon
energy larger than the energy gap.
% note 1
Using i.e. a single weak continuous linearly polarized laser beam, is possible
to promote electrons in $\mathbf{k}$ space regardless the symmetry of the system
resulting in a net current equal to zero. Then, if the system presents inversion
of symmetry, electrons promoted to the conduction bands at opposite $\mathbf{k}$
points will have opposite spin polarization resulting in a total spin injection
equal to zero. If the phenomena is produced in a noncentrosymmetric
semiconducting media the spin polarization injected at a given $\mathbf{k}$
point can be held\cite{alvaradoPRL85}, resulting in a PSC because the velocity
of electrons at opposite $\mathbf{k}$ points are in opposite directions.

\subsection{Spin velocity injection} % (fold)
\label{sec:theory-pure_spin_current}

We define the SVI as the velocity at which the spin, polarized along the
direction $\mathrm{a}$, propagates along the direction $\mathrm{b}$ as
\begin{equation}\label{eq:vab-w}
\mathcal{V}^{\mathrm{ab}}(\omega) \equiv
\frac{\dot{K}^{\mathrm{ab}}(\omega)}{(\hbar/2) \dot{n}(\omega)},
\end{equation}  
where the pure spin density injection current, $\dot{K}^{\mathrm{ab}} (\omega)$,
and the carrier injection rate, $\dot{n}(\omega)$, are given by
% 
\begin{equation}\label{eq:dot-kn}
\begin{aligned}
\dot{K}^{\mathrm{ab}}(\omega) 
&=
\mu^{\mathrm{abcd}}(\omega)
E^{\mathrm{c}}(\omega) E^{\mathrm{d*}}(\omega),
% \label{eq:dotk} 
\\
\dot{n}(\omega)
&=
\xi^{\mathrm{ab}}(\omega) E^{c }(\omega) E^{d*}(\omega),
% \label{eq:dotn}
\end{aligned}
\end{equation}
where the roman superscripts denote Cartesian directions, and if repeated, they
are to be summed over, the $\xi^{\mathrm{ab}}(\omega)$ are the carrier
generation rate tensor components and $\mu^{\mathrm{abcd}}(\omega)$ are the pure
spin-current pseudotensor components given by\cite{bhatPRL05}
\begin{equation*}\label{eq:mu}
\begin{aligned}
\mu^{\mathrm{abcd}}  (\omega) &
=
\frac{\pi e^{2}}{\hbar^{2}} \int 
\frac{d^{3}K}{8 \pi^{3}} \times \\
\sum_{vcc'}^{'}
\mathrm{Re} & \left[ K^{\mathrm{ab}}_{cc'} 
\left( 
r^{\mathrm{c}}_{vc'} 
r^{\mathrm{d}}_{cv } +
(c \leftrightarrow d)
\right) 
\right] \delta(\omega-\omega_{cv}),
\end{aligned}
\end{equation*}
where $K^{\mathrm{ab}}_{mn}(\mathbf{k}) =
\sum_{\ell}v^{\mathrm{a}}_{nl}(\mathbf{k}) S^{\mathrm{b}}_{lm}(\mathbf{k})$ are
the spin current matrix elements that, using time reversal invariance, satisfy
the relation $K^{\mathrm{ab}}_{nm}(\mathbf{-k}) = K^{\mathrm{ab*}}(\mathbf{k})
$. 
% 
The $'$ in the sum means that $c$ and $c'$ are quasi degenerate states and the
sum only covers these states and since $\mu^{\mathrm{abcd}}(\omega)$ is real we
have that $\mu^{\mathrm{abcd}}(\omega) = \mu^{\mathrm{abdc}}(\omega)$.
% 
Since we have 2D structures we use an incoming electric field parallel to the
surface, $\mathbf{E}^{a}(\omega) = E^{\mathrm{a}} e^{i(\alpha + \omega t)}$,
where the angle $\alpha$ corresponds to the linear polarization angle measured
positively in the counter-clockwise from the $x$ direction on the
surface of the structures.
% 
Then from Eq. \eqref{eq:dot-kn} we can rewrite the Eq. \eqref{eq:vab-w} and 
including the polarization angle dependence we have
\begin{widetext}
\begin{align}
\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)
&= \frac{2}{\hbar}
\frac{\mu^{\mathrm{abxx}}(\omega)
E^{2}(\omega)\cos^{2}(\alpha) + 
\mu^{\mathrm{abyy}}(\omega)
E^{2}(\omega)\sin^{2}(\alpha) + 
2\mu^{\mathrm{abxy}}(\omega)
E^{2}(\omega)\cos(\alpha)\sin(\alpha)}
{\xi^{\mathrm{xx}}(\omega)
E^{2}(\omega)\cos^{2}(\alpha) + 
\xi^{\mathrm{yy}}(\omega)
E^{2}(\omega)\sin^{2}(\alpha)},
\nonumber \\
&= \frac{2}{\hbar}
\frac{\mu^{\mathrm{abxx}}(\omega)\cos^{2}(\alpha) + 
\mu^{\mathrm{abyy}}(\omega)\sin^{2}(\alpha) + 
\mu^{\mathrm{abxy}}(\omega)\sin(2\alpha)}
{\xi^{\mathrm{xx}}(\omega)\cos^{2}(\alpha) + 
\xi^{\mathrm{yy}}(\omega)\sin^{2}(\alpha)},
\label{eq:vab-aw}
\end{align}
\end{widetext}
% 
If the polarization angle is fixed to $\alpha = \frac{\pi}{4}$ the previous
expression can be reduced to 
\begin{equation*}\label{eq:vab-90deg}
\mathcal{V}^{\mathrm{ab}} (\omega)
= \frac{2}{\hbar}
\frac{\mu^{\mathrm{abxx}}(\omega) + \mu^{\mathrm{abyy}}(\omega) + 
2\mu^{\mathrm{abxy}}(\omega)}
{\xi^{\mathrm{xx}}(\omega) + \xi^{\mathrm{yy}}(\omega)}.
\end{equation*}
Two interesting possibilities to analyze the SVI are fixing the spin along any
of the three Cartesian coordinates and, for our cases, particularly for the spin
directed along $z$ directed perpendicularly to the surface of the structure; or
fixing the velocity along the $x$ or $y$ Cartesian coordinates on the $xy$ plane
of the structures. In following subsections we present these cases.

@@

Because we have 2D structures, one of the options to analyze the spin velocity
(Eq. \eqref{eq:vab-aw}) is fixing the spin parallel to the surface of the
structure in the $x$ or $y$ directions or perpendicularly to it in the $z$
direction. We define the magnitude of the spin velocity with spin polarization
along the $\mathrm{b}$ Cartesian coordinate as
\begin{equation}
|\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)| 
=
\sqrt{
[\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)]^{2}\ +
[\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)]^{2}\ 
}, 
\label{eq:vs-mag}
\end{equation}
and the angle at which the spin velocity is directed on the $xy$ plane as
\begin{equation}
\gamma_{\mathrm{b}} (\omega,\alpha)
=
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)} \right),
\label{eq:gamma-ang}
\end{equation}
where the angle is measured in the counter-clockwise direction from the
positive $x$ Cartesian coordinate. {\color{red} We also define two special
angles
\begin{align}
\gamma_{\mathrm{b \parallel}}(\omega,\alpha) &= \alpha, 
\label{eq:gamma-par} 
\\
\gamma_{\mathrm{b \perp}}(\omega,\alpha) &= \alpha \pm 90^{\circ}.
\label{eq:gamma-perp}
\end{align}
The first corresponds to the case when the spin velocity is directed, on the
$xy$ plane in the same direction of the polarization angle of the incoming
beam; the second one corresponds to the case when the spin velocity is directed
perpendicularly respect to the polarization angle of the incoming beam. }

@@

{\color{red} Another possibility to analyze the Eq. \eqref{eq:vab-aw} is fixing
the velocity on the $xy$ plane along $x$ or $y$ Cartesian coordinate and then
define the magnitude of the spin velocity directed along the $\mathrm{a}$
direction as}
\begin{align}
|\mathcal{V}^{\mathrm{a}}&(\omega,\alpha)| = \nonumber \\
&\sqrt { 
[\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{az}}(\omega,\alpha)]^{2} .
},
\label{eq:vv-mag}
\end{align}
{\color{red} Then, the spin direction depends of the $x$, $y$, and $z$
components of the previous equation and so we define the spin orientation polar
and azimuthal angles as}
\begin{align}
\theta_{\mathrm{a}}  (\omega,\alpha)
=& 
\cos^{-1} \left( \frac{\mathcal{V}^{\mathrm{az}}(\omega,\alpha)}
{|\mathcal{V}^{\mathrm{a}}(\omega,\alpha)|} \right),
& 0 \leq &\theta \leq \pi, 
\label{eq:polar-ang}
\\
\varphi_{\mathrm{a}} (\omega,\alpha)
=& 
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)} \right),
& 0 \leq &\varphi \leq 2\pi.
\label{eq:azimuthal-ang} 
\end{align}
{\color{red} where $\theta_{\mathrm{a}}(\omega,\alpha)$ is measured from the
positive to the negative $z$ Cartesian coordinate and
$\varphi_{\mathrm{a}}(\omega,\alpha)$ is measured on the $xy$ plane in the
counter-clockwise direction from the positive $x$ Cartesian coordinate.}

@@

\subsection{Spin velocity injection} % (fold)
\label{sec:res-spin_velocity}

@@

{\color{red}Using the Eq. \eqref{eq:vab-aw}, we calculated the
$\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ response for the \emph{up} and
\emph{alt} 2D structures and for the CdSe and GaAs bulk systems; the results
are presented in Fig. \ref{fig:vab-str-comp}. The angle $\alpha$ presented in
the response of each structure is that for which the response is maximized in
each case.}
% 
From the figure we have that the onset of the response starts when the energy
of the incoming beam is the same of the gap energy.
% 
The most intense response corresponds to the \emph{up} structure centered at
0.088\,eV corresponding to the {\color{red} Mid Infrared (MIR)} radiation and
reaching a spin velocity of 87.2\,Km/s.
% 
In the other hand, for an energy range from 0.66\,eV to 3.0\,eV, corresponding
to energies of the {\color{red} Near Infrared (NIR) to visible radiation}, all
the four structures have contributions in the same order of magnitude.
% 
For this energy range the \emph{up} structure has two peaks centered at
1.94\,eV and 1.97\,eV reaching spin velocities of 22.2\,Km/s and -29.7\,Km/s,
respectively, and the \emph{alt} structure has two peaks centered at 0.72\,eV
and 0.91\,eV reaching spin velocities of -40.2\,Km/s and -32.9\,Km/s,
respectively.
% 
Then, for the bulk structures we have that the CdSe has only one intense
response centered at 0.91\,eV reaching a spin velocity of -26.9\,Km/s, and the
GaAs structure has a large and almost planar zone where the response is held
reaching the maximum for an incoming beam of energy of 2.31\,eV and resulting
in a spin velocity of -21.6\,Km/s.
% 
{\color{red} A negative quantity in the spin velocity means a change in the
spin polarization traveling in the opposite direction.
% 
In table \ref{tab:vab-str-comp} we present the comparison of this values for
the 2D and bulk structures. We found that the most intense response for the
spin velocity corresponds to the \emph{up} structure being 3.25 times more
intense than that of the CdSe and 4.03 times more intense than that of the GaAs
bulk structures. Also, the \emph{alt} structure has a response more intense
than the bulk systems but being less intense than the corresponding to the
\emph{up} one.}
%
